### 2016-12-01 paper.tex
initial

% Created 2016-11-30 Wed 17:09
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\author{Max Curran and Nick Merrill}
\date{\today}
\title{One-step, three-factor authentication with in-ear EEG}
\hypersetup{
 pdfauthor={Max Curran and Nick Merrill},
 pdftitle={One-step, three-factor authentication with in-ear EEG},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.0.50.2 (Org mode 9.0.1)}, 
 pdflang={English}}
\begin{document}

\maketitle
\begin{abstract}
We propose the first research study of one-step three-factor authentication.
Specifically we seek to demonstrate its feasibility and quantify its performance
using Ear EEG (electroencephalogram). By performing a single mental task, our
user will be able to present three authenticators at once – a knowledge factor
(their chosen secret thought and/or mental task), an inherence factor (their
brainwave signals), and a possession factor (the EEG sensing earpiece that is
custom-fitted to their ear). We will build the custom-fit Ear EEG earpieces and
conduct an experimental study to evaluate the accuracy and usability of this
authentication method.
\end{abstract}
\section{Introduction and Motivation}
\label{sec:orgd88076b}

It is well appreciated by experts and end-users alike that strong authentication is
critical to cybersecurity and privacy, now and into the future. Unfortunately,
news reports of celebrity account hackings serve as regular reminders that
the currently dominant method of authentication in consumer applications, 
single-factor authentication using passwords or other user-chosen secrets, 
is faced by many challenges. Major industry players such as Google and
Facebook have been strongly encouraging their users to adopt two-factor
authentication (2FA). However, the need for users to submit two different 
authenticators in two separate steps has frustrated wide adoption, 
due its additional hassle cost to the users. For instance, the popular Apple
iPhone has already implemented the necessary technologies to support device
unlock using either a user-selected passcode or a fingerprint. Therefore the
device could easily support a two-step two-factor authentication scheme if
desired. However, it is easy to understand why users would balk at having to
enter a passcode \emph{and} provide a fingerprint each time they want to unlock their phone.

In our previous work, we have proposed “one-step two-factor authentication” as a
new approach to authentication that can provide the security benefits of two-
factor authentication without incurring the hassle costs of two-step verification [1].
By employing consumer-grade EEG (electroencephalogram) sensing
technologies, we demonstrated in our 2013 passthoughts study that a user can
submit both a knowledge factor (i.e., secret thought) and an inherence factor
(i.e., brainwave signal unique to the individual) in a single step by performing a
single mental task [2]. We demonstrated the robustness of this method against
impersonation attacks, including conditions where the attacker may have learned
the target’s secret thought and/or secret task [3].

In the present proposal, we will undertake, to the best of our knowledge, the first
ever study of one-step three-factor authentication. In computer security,
authenticators are classified into three types: knowledge factors (e.g., passwords
and PINs), possession factors (e.g., physical tokens, ATM cards), and inherence
factors (e.g., fingerprints and other biometrics). Because three-factor
authentication (3FA) requires the user to submit one distinct instance of each
type of authenticator, it represents the strongest level of authentication security
possible.

We propose the use of custom-fit Ear EEG technology as the platform for
investigating the feasibility, performance, and usability of one-step three-factor
authentication. In addition to the same knowledge factor and inherence factor as
in our previous work, the user can submit in the same step the possession factor
in the form of the EEG-sensing ear-piece(s) that are custom-fitted to and worn in
their ear. These earpieces can serve as physical tokens in the same way as bank
ATM cards and wearable hardware tokens. Furthermore, because the earpieces
are custom-fitted to each individual, they will not be able to produce good
electrical impedances when worn by a different individual.

\section{Related Work}
\label{sec:orge0e46dc}
The use of EEG as a biometric signal for user authentication has a short history.
In 2005, Thorpe et al. motivate and outline the design of a passthoughts system,
where users authenticate by thinking of a passthought rather than typing in a
password [6]. Since 2002, a number of independent groups have achieved 99-
100\% authentication accuracy using multi-channel sensors placed on the scalp
[7-10]. In 2013, our group showed that 99\% authentication accuracy can also be
achieved using a consumer-grade single-channel sensor [2]. In particular, the
lack of signal diversity from multiple EEG channels can be overcome by allowing
the users to choose their own personalized passthoughts (e.g., sing their favorite
song in their head). There are two significant consequences of this result. First,
the passthoughts approach is no longer constrained by the high cost (\textasciitilde{}\$10k’s)
and low usability (gel-based electrodes; aesthetic challenges of an EEG cap) of
medical-grade multi-channel devices. Second, because users can choose and
easily change their secret mental task, this approach can support one-step two-
factor authentication [1] via the simultaneous presentation of the inherence factor
(brainwave signatures due to the unique folding structures of the cortex) and the
knowledge factor (the secret mental task).

Research in in-ear EEG is only several years old. Nonetheless, the concept has
attracted a lot of attention because of the discreetness factor of in-ear EEG over
traditional scalp-based EEG. A research team at the Imperial College London
and Aarhus University published a landmark paper in 2011 that introduced the
concept of in-ear EEG, demonstrating for the first time the feasibility of recording
brainwave signals from within the ear canal [11]. Follow-up work from the same
group demonstrated its ability to produce signal-to-noise ratios comparable to
those from conventional EEG electrode placements, robustness to common
sources of artifacts, and use in a brain-computer interface (BCI) system based on
auditory evoked potentials and visual evoked potentials [12-14]. Our 2016 study
[4] was the first to merge these two streams of work, using in-ear EEG signals for
user authentication. United Sciences is currently developing a consumer
hearable called The Aware that will measure EEG from the ear [15].
Behavioral authentication methods such as keystroke dynamics [16] and speaker
authentication [17] can be categorized as one-step two-factor authentication
schemes. In both cases, the knowledge factor (password or passphrase) and
inherence factor (typing rhythm or speaker’s voice) are employed. In contrast, the
Nymi band [18] supports one-step two-factor authentication via the inherence
factor (cardiac rhythm that is supposed to be unique to each individual) and the
possession factor (the wearing of the band on the wrist). However, as far as we
know, no one has proposed or demonstrated a one-step three-factor
authentication scheme.

\section{Proposed project}
\label{sec:org6246122}

We propose the first ever research study on one-step three-factor authentication.
The study will yield a number of novel contributions to the literature. First, we
expect to demonstrate the feasibility of presenting three authenticators in a single
step. Second, we will quantify the performance accuracy of user authentication
using custom-fit in-ear EEG hardware. Third, we can quantify the relative
performance of signals captured from different locations within the ear. Fourth,
we will evaluate the effectiveness of new classes of mental tasks for EEG-based
authentication. Fifth, we will evaluate the usability of custom-fit in-ear EEG
hardware.

\section{Research plan}
\label{sec:org5e36c1a}

There will be four key components and major research tasks to this proposed
study: (i) design and build custom-fit hardware, (ii) design and implement
authentication tasks, (iii) experimental platform and protocol for data collection,
and (iv) analysis of authentication performance.

In our original 2013 passthoughts study [2], we used unmodified commercial off-
the-shelf EEG sensing hardware. In our 2016 Ear-EEG passthoughts study [4],
we modified the off-the-shelf hardware in order to support data collection from
within the ear canal. Once modified, the same device could be used for all
participants. In the current study, we propose to build in-ear EEG sensors that
are custom-fit to the contours of the ears and ear canals of each individual
participant. Therefore, it is necessary for us to build a separate pair of devices for
each participant. In partnership with our research collaborators at Starkey
Hearing Science, who specialize in hearing-aid technologies, we will acquire
moldings of participants’ ears, and use the molds to build custom-fit acrylic
earpieces with embedded EEG electrodes.

A key research challenge is to design and build the electrodes to produce signals
of robust quality, given the small volume of the ear cavity, the small surface
areas of the electrodes, and the limited distances between the electrodes within
the ear. We will need to experiment with the locations of the electrodes, including
the reference and ground electrodes, which may be placed on the ear lobes or
behind the ears.

We propose an update and expansion of the authentication tasks, to take
advantage of our lessons learned from previous studies on the relative strengths
of different mental tasks, with regards to authentication accuracy and usability as
reported by participants. Furthermore, given the in-ear placement of the
electrodes, and hence spatial proximity to the temporal lobes, we see a unique
opportunity to design and test novel authentication tasks based on either auditory
imagery or auditory steady-state response (ASSR). We will design a suite of
authentication tasks to include:

\begin{itemize}
\item Baseline tasks
\item Motor imagery tasks (with personal secret)
\item Visual imagery tasks (with personal secret)
\item Audio imagery tasks (with personal secret)
\item Mixed audio/visual imagery tasks (with personal secret)
\item Auditory steady-state response (ASSR) tasks (using external stimuli)
\end{itemize}

This will allow us to quantify the security and usability performances of these
tasks, both collectively as they contribute to overall authentication accuracy, as
well as comparatively against one another.
We will use OpenBCI, a open-source EEG biosensing system as our data
collection platform. This requires integrating the custom-fit Ear EEG devices with
the OpenBCI board. We will also need to extend Indra, our own data collector
software, to support the recording of OpenBCI data that is synchronized with
PsychoPy, which we use for stimuli presentation. We will recruit on the order of
7-10 participants who are willing to commit to multiple session visits, first for ear
molding, then for device impedance testing, and then for one or more data
collection sessions.

For authentication analysis, we will build upon the analysis methodology and
tools from our previous studies. We will apply the logarithmic binning technique
that we have developed in [5] to the EEG power spectrum data during the signal
pre-processing step. We will then apply machine learning to train and test a
support vector classifier as in [4]. We will compare the results against the
threshold-based authentication protocol based on cosine similarity as in [2,3].
A key change for this study is that we will be able to collect up to 8 separate
channels of EEG data, including multiple channels from each ear. This is in
contrast to our previous studies, where we collected only a single channel of
EEG, either on the prefrontal cortex (FP1) location or in the ear canal. Therefore,
we will need to develop new analytic techniques to both (i) assess improvements
in authentication accuracy due to the increased channels of data, and (ii) assess
relative contributions of individual electrode locations to authentication accuracy.
Finally, we will evaluate the usability of custom-fit Ear EEG authentication along
several dimensions. They include the comfort and fit of the earpieces, the
preparatory steps of using the devices, the ease and repeatability of the
authentication tasks, and the recall rates of the personal secrets.
\section{{\bfseries\sffamily TODO} Preliminary results}
\label{sec:orgb471f1b}

We performed a very small pilot study (n=2) to better understand how feasible 
our project would be. Our cursory analysis indicates that we can successfully
detect EEG signals from the ear (Section 5.1), that these signals can be used
for reliable authentication among two people (Section 5.2), and that the success
of the authenticator relies on the user's chosen secret, and not just the inherence
factor of the user's unique signal (Section 5.3).

\subsection{{\bfseries\sffamily TODO} Validation}
\label{sec:orgdae54ad}
\subsection{{\bfseries\sffamily TODO} Authentication performance}
\label{sec:org14de4b9}
\subsection{Disentangling inherence and knowledge}
\label{sec:org91c37b4}

Though our classifier's accuracy is strong, we still cannot say that it relies 
on both knowledge and inherence to authenticate the subject. Is it the passthought
that the classifier is authenticating, or do users (or earphones) simply have characteristic
EEG readings?

If our classifier \emph{only} relied on inherence, we would expect the right person performing
the wrong task to reliably authenticate with the classifier described in Section 5.2.
In other words, we would expect the FAR for right-person-wrong-task to be the same 
as the FAR for right-person, right-task. We calculated the FAR for each task, and performed
a two-tailed t-test to determine our confidence that this set of FARs was drawn from
from a different distribution from those generated during the right-person, right-task trials.

\begin{center}
\begin{tabular}{lrrr}
Task & Right ear & Left ear & fp1\\
\hline
speech & 0.311 & 0.116 & 0\\
face & 0.355 & 0.055 & 0\\
breathe & 0 & 0.240 & 0\\
listennoise & 0.322 & 0.078 & 0\\
song\(_{\text{o}}\) & 0 & 0.166 & 0\\
song & 0 & 0.211 & 0\\
sequence & 0.433 & 0 & 0\\
listentone & 0 & 0.333 & 0\\
breathe\(_{\text{o}}\) & 0.254 & 0.129 & 0\\
sport & 0.222 & 0.135 & 0\\
\hline
\textbf{p-value} & \textbf{*0.0076} & \textbf{*0.009} & N/A\\
\end{tabular}
\end{center}
Our low p-values (Table 3) indicate that the passthought, as well as the inherence
factor associated with the custom-fit EEG earbud, both contribute to the classifier
performance discussed in Section 5.2. We are confident, then, in our claim that this work
could potentially yield multiple-factor authentication in a single step. However, we
will need to collect a great deal more data to substantiate this claim rigorously.

\section{Schedule and Budget}
\label{sec:org1590dbf}
We propose a 12-month schedule with two graduate students working on the
project:

\begin{itemize}
\item Months 1-6: hardware development for custom-fit earpiece; integration with OpenBCI; testing
\item Months 1-6: software integration between OpenBCI, PsychoPy, and Indra; design of new authentication tasks and implementation in PsychoPy
\item Months 7-8: experimental data collection
\item Months 9-12: data analysis; write and submit research papers
\end{itemize}
\section{References}
\label{sec:org14ba974}

[1]
J. Chuang, One-Step Two-Factor Authentication with Wearable Bio-Sensors, Workshop on
"Who are you?! Adventures in Authentication" (WAY'14), 10th USENIX Symposium on
Usable Privacy and Security (SOUPS'14), 2014.

[2]
J. Chuang, H. Nguyen, C. Wang, B. Johnson, I Think, Therefore I Am: Usability and
Security of Authentication Using Brainwaves, Workshop on Usable Security (USEC'13),
17th Int. Conference on Financial Cryptography and Data Security (FC’13), 2013.

[3]
B. Johnson, T. Maillart, J. Chuang, My Thoughts are Not Your Thoughts: Robustness of
Brainwave Signal Authentication Against Impersonation Attacks, Workshop on Usable
Privacy \& Security for wearable and domestic Ubiquitous Devices, ACM UbiComp, 2014.

[4]
M. Curran, J. Yang, N. Merrill, J. Chuang. Passthoughts Authentication with Low Cost
EarEEG. Proc. 38th Intl. Conf. of IEEE Engineering in Medicine and Biology Society
(EMBC 2016), Aug. 2016.

[5]
N. Merrill, Maillart, Johnson, J. Chuang, Improving Physiological Signal Classification Using
Logarithmic Quantization and a Progressive Calibration Technique, Proc. 2nd Intl. Conf. on
Physiological Computing Systems (PhyCS'15), Feb. 2015.

[6]
J. Thorpe, P. van Oorschot, and A. Somayaji. Pass-thoughts: Authenticating with our
minds. In Proceedings of the New Security Paradigms Workshop (NSPW), 2005.

[7]
M. Poulos, M. Rangoussi, N. Alexandris, and A. Evangelou. Person identification from the
EEG using nonlinear signal classification. Methods of Information in Medicine, 2002.

[8]
S. Marcel and J. del R. Millan. Person authentication using brainwaves (EEG) and
maximum a posteriori model adaptation. IEEE Transactions on Pattern Analysis and
Machine Intelligence, 29(4), April 2007.

[9]
R. Palaniappan. Two-stage biometric authentication method using thought activity brain
waves. International Journal of Neural Systems, 18(1):59–66, 2008.

[10]
C. Ashby, A. Bhatia, F. Tenore, and J. Vogelstein. Low-cost electroencephalogram (EEG)
based authentication. In Proc. of 5th Int. IEEE EMBS Conf. on Neural Engineering, 2011.

[11]
D. Looney, C. Park, P. Kidmose, M. L. Rank, M. Ungstrup, K. Rosenkranz, and D. P.
Mandic, “An in-the-ear platform for recording electroencephalogram,” in Proc. Int. Conf.
IEEE Eng. Med. Biol. Soc., 2011, pp. 6682–6885.

[12]
D. Looney, P. Kidmose, C. Park, M. Ungstrup, M. Rank, K. Rosenkranz, and D. Mandic,
“The in-the-ear recording concept,” IEEE Pulse, vol. 3, no. 6, pp. 32–42, Nov./Dec. 2012.

[13]
P. Kidmose, D. Looney, M. Ungstrup, M.L. Rank, D.P. Mandic, “A study of evoked
potentials from ear-EEG,” IEEE Trans. Biomed. Eng. 60(10), pp. 2824–2830, 2013.

[14]
P. Kidmose, D. Looney, and D. P. Mandic, “Ear-EEG from generic earpieces: a feasibility
study,” in Proc. Int. Conf. IEEE Eng. Med. Biol. Soc., 2013, pp. 543-546.

[15] The Aware. \url{http://efitaware.com} [Online; accessed 22-Oct-2016].

[16] F. Monrose and A. Rubin. Authentication via keystroke dynamics. In Proceedings of the 4th
ACM conference on Computer and communications security, pages 48–56. ACM, 1997.

[17]
K. Stevens et al. Speaker authentication and identification: a comparison of spectrographic
and auditory presentations of speech material. The Journal of the Acoustical Society of
America 44.6 (1968): 1596-1607.

[18] Nymi. \url{https://nymi.com} [Online; accessed 22-Oct-2016].

[19] E. Strickland. In-Ear EEG Makes Unobtrusive Brain-Hacking Gadgets a Real Possibility.
IEEE Spectrum, 7/7/2016.

[20]
N. Merrill, M. Curran, J. Yang, J. Chuang. Classifying Mental Gestures with In-Ear EEG.
Proc. 13th IEEE Intl. Conf. on Wearable and Implantable Body Sensor Networks (BSN
2016), Jun. 2016.

[21]
E. Sedenberg, J. Chuang, D. Mulligan. Designing Commercial Therapeutic Robots for
Privacy Preserving Systems and Ethical Research Practices Within the Home. Intl. J. of
Social Robotics, Aug. 2016.

[22] E. Sedenberg et al. A Window into the Soul: Biosensing in Public. Under review, 2017.

[23] R. Wong et al. “All that happens must be known”: Eliciting Values and Conceptions of
Privacy using Science Fiction. Under review, 2017.

[24]
R. Wong et al. Real-Imaginary Entanglements: Using Science Fiction and Design Fiction to
Explore and Question Sensing and Tracking Technologies. Under review, 2017.
\end{document}